---
title: "Food Module: Food Factors"

author: |
  | **Bruno Caetano Vidigal**
  | Food and Agriculture Organization of the United Nations
output: pdf_document
---
---

```{r include=FALSE, cache=FALSE}
## Load required functions
library(data.table)
library(ggplot2)
library(scales)
library(knitr)
library(devtools)
#install_github("ndphillips/yarrr")
library("yarrr")
library(RColorBrewer)
library(data.table)
library(faosws)
library(dplyr)
library(faoswsUtil)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE, cache=FALSE}
## set up for the test environment and parameters
# R_SWS_SHARE_PATH <- Sys.getenv("R_SWS_SHARE_PATH")
# DEBUG_MODE <- Sys.getenv("R_DEBUG_MODE")
# overwritableFlags = c("M", "I")

# This return FALSE if on the Statistical Working System
if(CheckDebug()){
    
    message("Not on server, so setting up environment...")
    
    library(faoswsModules)
    SETTINGS <- ReadSettings("C:/Users/caetano/Documents/Github/faoswsFood/modules/impute_food/sws.yml")
    
    # If you're not on the system, your settings will overwrite any others
    R_SWS_SHARE_PATH <- SETTINGS[["share"]]
    
    # Define where your certificates are stored
    SetClientFiles(SETTINGS[["certdir"]])
    
    # Get session information from SWS. Token must be obtained from web interface
    GetTestEnvironment(baseUrl = SETTINGS[["server"]],
                       token = SETTINGS[["token"]])
    files = dir("~/Github/faoswsFood/R", full.names = TRUE)
    sapply(files, source)
    
}
```

```{r include=FALSE, cache=FALSE}
# Food Factors

areaCodesFoodFactors <- GetCodeList("food", "food_factors","geographicAreaM49")$code
comCodesFoodFactors <- GetCodeList("food", "food_factors","foodCommodityM")$code
fdmCodesFoodFactors <- GetCodeList("food", "food_factors","foodFdm")$code
funCodesFoodFactors <- GetCodeList("food", "food_factors","foodFunction")$code
varCodesFoodFactors <- "y_e" ## Only need elasticities from the food domain table

## Define the dimensions
dimM49FoodFactors <- Dimension(name = "geographicAreaM49", keys = areaCodesFoodFactors)
dimComFoodFactors <- Dimension(name = "foodCommodityM", keys = comCodesFoodFactors)
dimFdmFoodFactors <- Dimension(name = "foodFdm", keys = fdmCodesFoodFactors)
dimFunFoodFactors <- Dimension(name = "foodFunction", keys = funCodesFoodFactors)
dimVarFoodFactors <- Dimension(name = "foodVariable", keys = varCodesFoodFactors)

## Define the keys
keyFdmFoodFactors <- DatasetKey(domain = "food", dataset = "food_factors",
                     dimensions = list(dimM49FoodFactors, dimComFoodFactors, 
                                       dimFdmFoodFactors, dimFunFoodFactors, 
                                       dimVarFoodFactors))

## download the food dimension data (elasticities) from the SWS
foodFactors <- GetData(keyFdmFoodFactors, flags=FALSE, normalized = FALSE)
foodFactors <- nameData("food", "food_factors", foodFactors)
setnames(foodFactors, "Value_foodVariable_y_e", "elasticity")
setnames(foodFactors, "foodFdm", "foodDemand")
setnames(foodFactors, "foodCommodityM", "foodCommodity")
foodFactors[, c("foodFdm_description", "foodFunction_description") := NULL]
```


```{r include=FALSE, cache=FALSE}

# Food Factors Nat
areaCodesFoodFactorsNat <- GetCodeList("food", "food_factors_n","geographicAreaM49")$code
comCodesFoodFactorsNat <- GetCodeList("food", "food_factors_n","foodCommodityM")$code
fdmCodesFoodFactorsNat <- GetCodeList("food", "food_factors_n","foodFdm")$code
funCodesFoodFactorsNat <- GetCodeList("food", "food_factors_n","foodFunction")$code
varCodesFoodFactorsNat <- "y_e" ## Only need elasticities from the food domain table

## Define the dimensions
dimM49FoodFactorsNat <- Dimension(name = "geographicAreaM49", keys = areaCodesFoodFactorsNat)
dimComFoodFactorsNat <- Dimension(name = "foodCommodityM", keys = comCodesFoodFactorsNat)
dimFdmFoodFactorsNat <- Dimension(name = "foodFdm", keys = fdmCodesFoodFactorsNat)
dimFunFoodFactorsNat <- Dimension(name = "foodFunction", keys = funCodesFoodFactorsNat)
dimVarFoodFactorsNat <- Dimension(name = "foodVariable", keys = varCodesFoodFactorsNat)



keyFdmNat <- DatasetKey(domain = "food", dataset = "food_factors_n",
                     dimensions = list(dimM49FoodFactorsNat, dimComFoodFactorsNat,
                                       dimFdmFoodFactorsNat, dimFunFoodFactorsNat,
                                       dimVarFoodFactorsNat))

foodFactorsNat <- GetData(keyFdmNat, flags=FALSE, normalized = FALSE)
foodFactorsNat <- nameData("food", "food_factors_n", foodFactorsNat)
setnames(foodFactorsNat, "Value_foodVariable_y_e", "elasticityNat")
setnames(foodFactorsNat, "foodFdm", "foodDemand")
setnames(foodFactorsNat, "foodCommodityM", "foodCommodity")
foodFactorsNat[, c("foodFdm_description", "foodFunction_description") := NULL]
```

## Abstract

This vignette provides a comparison between the datasets food_factors and food_factors_nat in the food domain.

# 1. Food Factors

```{r, echo=FALSE}
head(foodFactors, 4)
```


# 2. Food Factors Nat

```{r, echo=FALSE}
head(foodFactorsNat, 4)
```

# 3. Count

```{r, echo=FALSE}

foodFactorsTab <- foodFactors[, list(geographicAreaM49 = length(unique(geographicAreaM49)),
                   foodCommodity = length(unique(foodCommodity)),
                   foodDemand = length(unique(foodDemand)),
                   foodFunction = length(unique(foodFunction)),
                   nrows = nrow(foodFactors))]
foodFactorsTab <- foodFactorsTab[, dataset:= "food_factors"]
setcolorder(foodFactorsTab, c("dataset", "nrows", "geographicAreaM49", "foodCommodity", "foodDemand", "foodFunction"))

foodFactorsNatTab <- foodFactorsNat[, list(geographicAreaM49 = length(unique(geographicAreaM49)),
                      foodCommodity = length(unique(foodCommodity)),
                      foodDemand = length(unique(foodDemand)),
                      foodFunction = length(unique(foodFunction)),
                      nrows = nrow(foodFactorsNat))]
foodFactorsNatTab <- foodFactorsNatTab[, dataset:= "food_factors_nat"]
setcolorder(foodFactorsNatTab, c("dataset", "nrows", "geographicAreaM49", "foodCommodity", "foodDemand", "foodFunction"))

kable(rbind(foodFactorsTab, foodFactorsNatTab), 
      format = "markdown", padding = 0)

```


# 4. What does exist in food_factors and does not exist in food_factors_nat?

## 4.1 **geographicAreaM49** **Level**

```{r, echo=FALSE}

countryFoodFactors <- foodFactors[, .N, c("geographicAreaM49", "geographicAreaM49_description")]

countryFoodFactorsNat <- foodFactorsNat[, .N, c("geographicAreaM49", "geographicAreaM49_description")]
setnames(countryFoodFactorsNat, "geographicAreaM49_description", "geographicAreaM49_descriptionNat")

mergeCountries <- merge(countryFoodFactors, 
      countryFoodFactorsNat[, c("geographicAreaM49", "geographicAreaM49_descriptionNat"), with = F], 
      by="geographicAreaM49", all = T)

mergeCountries[is.na(geographicAreaM49_descriptionNat), .N, 
               c("geographicAreaM49", 
                 "geographicAreaM49_description")][, 
                                                      c("geographicAreaM49", 
                                                        "geographicAreaM49_description"),
                                                      with = F]

```


## 4.2 **foodCommodity** **Level**

```{r, echo=FALSE}

foodCommodityFoodFactors <- foodFactors[, .N, c("foodCommodity", "foodCommodityM_description")]

foodCommodityFoodFactorsNat <- foodFactorsNat[, .N, c("foodCommodity", "foodCommodityM_description")]
setnames(foodCommodityFoodFactorsNat, "foodCommodityM_description", "foodCommodityM_descriptionNat")

mergeCommodity <- merge(foodCommodityFoodFactors, 
                        foodCommodityFoodFactorsNat[, c("foodCommodity", "foodCommodityM_descriptionNat"), with = F], 
                        by="foodCommodity", all = T)


mergeCommodity[is.na(foodCommodityM_description), .N, 
               c("foodCommodity", 
                 "foodCommodityM_descriptionNat")][, 
                                                      c("foodCommodity", 
                                                        "foodCommodityM_descriptionNat"),
                                                      with = F]

```

# 5. What does exist in food_factors_nat and does not exist in food_factors?

## 5.1 **geographicAreaM49** **Level**

```{r, echo=FALSE}
mergeCountries[is.na(geographicAreaM49_description), .N, 
               c("geographicAreaM49", 
                 "geographicAreaM49_descriptionNat")][, 
                                                      c("geographicAreaM49", 
                                                        "geographicAreaM49_descriptionNat"),
                                                      with = F]
```

## 5.2 **foodCommodity** **Level**

```{r, echo=FALSE}

mergeCommodity[is.na(foodCommodityM_descriptionNat), .N, 
               c("foodCommodity", 
                 "foodCommodityM_description")][, 
                                                   c("foodCommodity", 
                                                     "foodCommodityM_description"),
                                                   with = F]

```
